---
title: "Fitting Three multistable displays"
output: html_document
---

This notebook fits history dependent Gamma distribution data for **Three Displays** data set, either for the original or simulated data (see _Select Set_ chunk). These include binocular rivalry (BR), kinectic-depth effect (KD), and Necker Cube (NC).


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(loo)
library(MASS)
library(patchwork)
library(purrr)
library(readr)
library(rstan)
library(tidyr)
```

```{r Tabula rasa}
rm(list = ls())
```

```{r}
file <- "3Displays"
taus <- c(0.9, 30)
```

## History choice

Select which history to use (`history_to_use` variable). By default, we used "Same" history, _i.e._, history computed for the percept that becomes dominant.

```{r history settings}
history_to_use <- "Same"
history_variants <- c("Same", "Other")
history_to_exclude <- setdiff(history_variants, c(history_to_use))
```

## Importing data

```{r import}
reports <- read_csv(fs::path("Data with History", stringr::str_c(file, ".csv")),
                    col_types = cols(Observer = col_character(),
                                     Group = col_character(),
                                     State = col_double(),
                                     Time = col_double(),
                                     Duration = col_double(),
                                     HistorySame = col_double(),
                                     HistoryOther = col_double(),
                                     Tau = col_double())) %>%
  
  filter(Tau %in% taus) %>%
  dplyr::select(-one_of(stringr::str_c("History", history_to_exclude))) %>%
  dplyr::rename(History = !!stringr::str_c("History", history_to_use))

# inverting same history, so that larger values mean lower adaptation levels
if (history_to_use == "Same")
  reports$History <- 1 - reports$History

# centering history based on its median value
reports <-
  reports %>%
  group_by(Tau, Observer) %>%
  mutate(cHistory = History - median(History, na.rm = TRUE))
```
  

## Gamma model

```{stan output.var="gamma_model"}
data{
  int<lower=1> rowN;          // number of rows in the table
  int<lower=1> participantsN; // number of participants
  int<lower=1> timescalesN;   // number of timescales

  int<lower=1> iParticipant[rowN];  // participant 

  real<lower=0> Duration[rowN]; // duration of the current percept
  matrix[rowN, timescalesN] History; // history before current percept onset
}
parameters{
  // --- intercepts, fully independent ---
  vector[participantsN] aShape;
  vector[participantsN] aScale;

  // --- main effect of the history, hierarchical ---
  // participants
  row_vector[timescalesN] bShape_History[participantsN];
  row_vector[timescalesN] bScale_History[participantsN];
  
  // group
  vector[timescalesN] bShape_History_Mu;
  vector<lower=0>[timescalesN] bShape_History_Sigma;
  vector[timescalesN] bScale_History_Mu;
  vector<lower=0>[timescalesN] bScale_History_Sigma;
}
transformed parameters{
  vector<lower=0>[rowN] shape;
  vector<lower=0>[rowN] scale;

  for(iRow in 1:rowN){
    shape[iRow] = exp(aShape[iParticipant[iRow]] + sum(bShape_History[iParticipant[iRow]] .* History[iRow]));
    scale[iRow] = exp(aScale[iParticipant[iRow]] + sum(bScale_History[iParticipant[iRow]] .* History[iRow]));
  }
}
model {
  // --- intercepts, fully independent ---
  aShape ~ normal(0, 10);
  aScale ~ normal(0, 10);

  // --- main effect of the history, population ---
  bShape_History_Mu ~ normal(0.0, 1.0);
  bShape_History_Sigma ~ cauchy(0, 10);
  bScale_History_Mu ~ normal(0, 1);
  bScale_History_Sigma ~ cauchy(0, 10);

  // parameters for each timescale
  for(iT in 1:timescalesN){
    // main effect of history, participants
    for(iP in 1:participantsN) {
      // main effect of history
      bShape_History[iP][iT] ~ normal(bShape_History_Mu[iT],
                                      bShape_History_Sigma[iT]);
      bScale_History[iP][iT] ~ normal(bScale_History_Mu[iT],
                                      bScale_History_Sigma[iT]);
     }
  }

  Duration ~ gamma(shape, inv(scale));
}
generated quantities{
  vector[rowN] log_lik;
  for (iRow in 1:rowN){
    log_lik[iRow] = gamma_lpdf(Duration[iRow] | shape[iRow], inv(scale[iRow]));
  }
}
```

## Fitting model for every group

Model is fitting for each group separately. Fits are save in "_Fits_" folder, posterior samples in "_Posterior samples_" folder.

```{r}
fit_display <- function(df){
  # putting history per tau in a separate column
  reports_wide <-
    df %>%
    mutate(Observer = as.factor(Observer), 
           iObserver = as.integer(Observer)) %>%
    dplyr::select(-History) %>%
    tidyr::pivot_wider(names_from = Tau, values_from = cHistory) %>%
    na.omit()
  
  # extracting history into separate matrix
  history_matrix <-
    reports_wide %>%
    select_at((ncol(.)-length(taus)+1) : ncol(.)) %>%
    as.matrix()
  
  data_for_stan <- list(
    rowN = nrow(reports_wide),
    participantsN = max(reports_wide$iObserver),
    timescalesN = ncol(history_matrix),
  
    iParticipant = reports_wide$iObserver,
    Participant = levels(reports_wide$Observer),
  
    Duration = reports_wide$Duration,
    History = history_matrix,
    
    Tau = taus,
    Group = reports_wide$Group[1]
  )
  
  model_fit <- rstan::sampling(gamma_model,
                               control=list(adapt_delta=0.98, max_treedepth=15),
                               include=FALSE,
                               pars=c("shape", "scale"),
                               cores = future::availableCores(),
                               data = data_for_stan)

  save(model_fit, file = glue::glue("Fits/{reports_wide$Group[1]}.Rdata"))

  posterior <- rstan::extract(model_fit)
  save(posterior,
       data_for_stan, 
       file = glue::glue("Posterior samples/{reports_wide$Group[1]}.Rdata"))

  get_bfmi(model_fit)

  tibble(divergent=  rstan::get_num_divergent(model_fit),
         max_treedepth =  rstan::get_num_max_treedepth(model_fit))
}

reports %>%
  group_by(Group) %>%
  do(fit_display(.))
```


